The Talking Heads Experiment

### 2015-04-27 chap9/chap9.tex
fixed remaining typos

The N01SE exhibition was curated by Adam Lowe (see \figref{fig:lowe}) and Simon Schaffer. 
Adam Lowe is an artist and technologist. He is currently the director of Factum arte (Madrid) which is specialised 
in making life-like replicas of paintings, sculptures and archeological objects, using laser-scanners and 3d printers. 
Recent realisations include the sculptures of Giambattista Piranesi \citep{Lowe:2010}, the 
paintings of Caravaggio and the tomb of Tutankhamun. 
Simon Schaffer is a science historian, professor at the University of Cambridge at the department of History and Philosophy 
of Science. He wrote extensively about the historical developments in scientific research \citep{Schaffer:2011}
and animated television and radio programs, including {\itshape Light Fantastic} (BBC4). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap9/figs/lowe.pdf}}
\caption{\label{fig:lowe}Adam Lowe (curator of N01SE) during installation of the Talking Heads experiment at the Wellcome Gallery in London. 
The backwall contained the figures. On the right wall, posters were displayed explaining the experiment.}
\end{figure}

The exhibition was anounced in the following way by its curators: 
\begin{quotation}
A multi-site multimedia exhibition in Cambridge with ``realtime" links to London, organised around three key themes in ``digitality'':

@@

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone elses N01SE.\todo{I have unset italics for this quotation}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour.pdf}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

New words accrue in two ways: either an agent creates its own new words from random combinations of syllables; or it stores transmitted words together with possible-meaning guesses inferred from the scene, then uses a hypothesis-test strategy to render lexicons mutually compatible. Agents keep a running score for every word-meaning pair in their lexicon. When word-topic recognition succeeds the score for that pair goes up, and that of other alternates goes down. This dynamic forces the lexicon of each individual agent toprogressively conform, and keeps it adapting to any language changes or new meanings that need to be expressed. During the course of the exhibition, a group of robotic agents autonomously constructs a shared language about real world scenes in front of them. Humans can interact with the installation through the Internet; they can teach their agents words and follow the general progress towards the construction of the language.

@@

The British hackers realised that they could give dirty names to their agents and, because they 
could teach their own agents these words, they could also teach their agents dirty words for colours, 
shapes, or any other concept that agents were using. These words would unavoidably propagate in the population. 
As these hackers were extremely active, creating many agents, continuously launching them to different sites, and teaching 
their agents words, the global vocabulary progressively became unacceptable. 
The installations were in public spaces visited by school children, and so concern grew with the exhibition organisers
at other locations (except the London site where those responsable for the exhibition
actively encouraged this destructive behaviour). 
As a response, the experiment was temporarily halted and provisions put in to avoid that a small group would have 
excessive influence. It was still possible to provide a name to your agent and teach your agents words but 
some form of decency had to be respected. This change restored order but resulted in an overreaction of the part of 
the English hacker group who now used all possible means to attack the Talking Heads servers themselves, 
enouraged by Denna Jones and aided by others at the Wellcome Gallery.

@@

A group of University of Hull students (with names like {\it Trash}, {\it yeah8a8y}, {\it schedski}) 
became very much involved. They 
communicated through the Forum using often slang and sexist language and making a surprising 
number of spelling errors and grammatical mistakes. The group had been 
trying to impose their own language by very actively teaching their agents and sending them to the same site, namely 
Paris, which they had noticed had the most stable operating conditions. This worked (as indeed it was supposed to) 
at which point they decided to do the same with the London site although the camera 
alignment was too unreliable to allow the evolution of a stable language.

@@

At this point, they realised that the London alignment was off and this is the first time that Phlox (as Denna 
Jones called herself) intervened. She immediately shows
a negative attitude towards the experiment triggered by her personal conflict with Adam Lowe. 
Colleagues of hers ({\it fish} and {\it A Londoner}) respond with inside joke
remarks that the Hull group does not understand. {\it Fish} had been trusted with admin passwords because he
was responsable for maintaining the London site, but he greatly abused this trust as the experiment proceeded.

@@

To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\it Phlox}) took 
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

Once the site came on-line, we began to remove the swearwords based on complaints from the public sites where the 
experiment was shown (in particular from Cambridge). This was the chance for Denna Jones (alias {\it phlox}) to stimulate 
attacks on the experiment's servers by playing on the sentiment of the players that there was a higher authority 
impinging on their freedom. The Hull hackers then started to divert the php script of the agents so that they 
could reschedule agents, encouraged by fish who had admin rights to the London site. However they became suspicious
of fish (who also had created an agent with the name Francis Crick)
because they realised he had these admin rights and therefore should (normally) have been part of the 
crew that maintained the overall system. Denna Jones tried to quell this suspicion by denying that he was an insider.

@@

The story continues as the Hull hackers entice Denna Jones ({\it Phlox}) to stand in front of one of 
the talking heads cameras so that they could see her. She actually does, to great acclaim of the hackers -- who 
seem surprised she is a woman. Denna Jones keeps further encouraging the Hull hackers. 
And from here on, the tone gets increasingly aggressive as loopholes are closed to prevent manipulation of the experiment
and attacks on the servers. There are again suspicions against {\it fish} who clearly has access from the inside using 
a password only given to trusted collaborators but which he abuses for destroying the experiment. The group is also now
beginning to communicate directly instead of through the {\it Forum} associated with the Talking Heads website.

@@

Why did we not interfere in what was going on or directly defend our points of view on the Forum? There were two reasons. First, our team was small and engaged with many other activities. The destructive activities of Denna Jones and her friends were not really worth our continuous attention and precious time. Second, this whole episode was yielding significant data about how individuals behave with respect to artificial agents and about the personality traits of those who are most likely to engage with them. The most obvious conclusion from these data is that we should not attempt to launch such experiments in 
the public domain, not because they are not feasible from a technical point of view, particularly today with much more 
reliable web technologies, but because the way that (some) individuals are likely 
to behave with respect to these technologies. The job of figuring out how to create artificial societies and cultures 
that can cooperate with human societies is an unsolved problems and it will require help from anthropologists to figure out 
how it can be set up. Undoubtly anonimity is one of the main sources of deviant behaviour \citep{Knight:1999}.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.30\textwidth]{chap9/figs/eyes.pdf}
\includegraphics[width=.30\textwidth]{chap9/figs/identified-colors.pdf}}
\caption{\label{fig:eyes}Example interaction of the ``Look into the Box" art piece. 
Left: Top and pixilated eye of a spectator. Right: the main colours that 
were extracted from this image. Agents played language games to describe these eye colours.}
\end{figure}

>>>>

The N01SE exhibition was curated by Adam Lowe (see \figref{fig:lowe}) and Simon Schaffer. 
Adam Lowe is an artist and technologist. He is currently the director of Factum arte (Madrid) which is specialised 
in making life-like replicas of paintings, sculptures and archaeological objects, using laser-scanners and 3d printers. 
Recent realisations include the sculptures of Giambattista Piranesi \citep{Lowe:2010}, the 
paintings of Caravaggio and the tomb of Tutankhamun. 
Simon Schaffer is a science historian, professor at the University of Cambridge at the department of History and Philosophy 
of Science. He wrote extensively about the historical developments in scientific research \citep{Schaffer:2011}
and animated television and radio programs, including {\itshape Light Fantastic} (BBC4). 
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.70\textwidth]{chap9/figs/lowe.pdf}}
\caption{\label{fig:lowe}Adam Lowe (curator of N01SE) during installation of the Talking Heads experiment at the Wellcome Gallery in London. 
The backwall contained the figures. On the right wall, posters were displayed explaining the experiment.}
\end{figure}

The exhibition was announced in the following way by its curators: 
\begin{quotation}
A multi-site multimedia exhibition in Cambridge with ``realtime" links to London, organised around three key themes in ``digitality'':

@@

N01SE celebrates the essential excess from which information is drawn. It probes many different ways of seeing and 
being in the world. Chances are your own sense of order is already someone else's N01SE.\todo{I have unset italics for this quotation}
\end{quotation}
%\begin{figure}[htbp]
%  \centerline{\includegraphics[width=.95\textwidth]{chap9/figs/latour.pdf}}
%\caption{\label{latour} 
%Luc Steels Discusses The N01SE exhibition with Bruno Latour who intensively followed the experiments already from the first series at the Laboratorium exhibition in Antwerp. His comments significantly improved the initial designs.}
%\end{figure}

@@

New words accrue in two ways: either an agent creates its own new words from random combinations of syllables; or it stores transmitted words together with possible-meaning guesses inferred from the scene, then uses a hypothesis-test strategy to render lexicons mutually compatible. Agents keep a running score for every word-meaning pair in their lexicon. When word-topic recognition succeeds the score for that pair goes up, and that of other alternates goes down. This dynamic forces the lexicon of each individual agent to progressively conform, and keeps it adapting to any language changes or new meanings that need to be expressed. During the course of the exhibition, a group of robotic agents autonomously constructs a shared language about real world scenes in front of them. Humans can interact with the installation through the Internet; they can teach their agents words and follow the general progress towards the construction of the language.

@@

The British hackers realised that they could give dirty names to their agents and, because they 
could teach their own agents these words, they could also teach their agents dirty words for colours, 
shapes, or any other concept that agents were using. These words would unavoidably propagate in the population. 
As these hackers were extremely active, creating many agents, continuously launching them to different sites, and teaching 
their agents words, the global vocabulary progressively became unacceptable. 
The installations were in public spaces visited by school children, and so concern grew with the exhibition organisers
at other locations (except the London site where those responsable for the exhibition
actively encouraged this destructive behaviour). 
As a response, the experiment was temporarily halted and provisions put in to avoid that a small group would have 
excessive influence. It was still possible to provide a name to your agent and teach your agents words but 
some form of decency had to be respected. This change restored order but resulted in an overreaction of the part of 
the English hacker group who now used all possible means to attack the Talking Heads servers themselves, 
encouraged by Denna Jones and aided by others at the Wellcome Gallery.

@@

A group of University of Hull students (with names like {\itshape Trash}, {\itshape yeah8a8y}, {\itshape schedski}) 
became very much involved. They 
communicated through the Forum using often slang and sexist language and making a surprising 
number of spelling errors and grammatical mistakes. The group had been 
trying to impose their own language by very actively teaching their agents and sending them to the same site, namely 
Paris, which they had noticed had the most stable operating conditions. This worked (as indeed it was supposed to) 
at which point they decided to do the same with the London site although the camera 
alignment was too unreliable to allow the evolution of a stable language.

@@

At this point, they realised that the London alignment was off and this is the first time that Phlox (as Denna 
Jones called herself) intervened. She immediately shows
a negative attitude towards the experiment triggered by her personal conflict with Adam Lowe. 
Colleagues of hers ({\itshape fish} and {\itshape A Londoner}) respond with inside joke
remarks that the Hull group does not understand. {\itshape Fish} had been trusted with admin passwords because he
was responsable for maintaining the London site, but he greatly abused this trust as the experiment proceeded.

@@

To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\itshape Phlox}) took 
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

Once the site came on-line, we began to remove the swearwords based on complaints from the public sites where the 
experiment was shown (in particular from Cambridge). This was the chance for Denna Jones (alias {\itshape phlox}) to stimulate 
attacks on the experiment's servers by playing on the sentiment of the players that there was a higher authority 
impinging on their freedom. The Hull hackers then started to divert the php script of the agents so that they 
could reschedule agents, encouraged by fish who had admin rights to the London site. However they became suspicious
of fish (who also had created an agent with the name Francis Crick)
because they realised he had these admin rights and therefore should (normally) have been part of the 
crew that maintained the overall system. Denna Jones tried to quell this suspicion by denying that he was an insider.

@@

The story continues as the Hull hackers entice Denna Jones ({\itshape Phlox}) to stand in front of one of 
the talking heads cameras so that they could see her. She actually does, to great acclaim of the hackers -- who 
seem surprised she is a woman. Denna Jones keeps further encouraging the Hull hackers. 
And from here on, the tone gets increasingly aggressive as loopholes are closed to prevent manipulation of the experiment
and attacks on the servers. There are again suspicions against {\itshape fish} who clearly has access from the inside using 
a password only given to trusted collaborators but which he abuses for destroying the experiment. The group is also now
beginning to communicate directly instead of through the {\itshape Forum} associated with the Talking Heads website.

@@

Why did we not interfere in what was going on or directly defend our points of view on the Forum? There were two reasons. First, our team was small and engaged with many other activities. The destructive activities of Denna Jones and her friends were not really worth our continuous attention and precious time. Second, this whole episode was yielding significant data about how individuals behave with respect to artificial agents and about the personality traits of those who are most likely to engage with them. The most obvious conclusion from these data is that we should not attempt to launch such experiments in 
the public domain, not because they are not feasible from a technical point of view, particularly today with much more 
reliable web technologies, but because the way that (some) individuals are likely 
to behave with respect to these technologies. The job of figuring out how to create artificial societies and cultures 
that can cooperate with human societies is an unsolved problems and it will require help from anthropologists to figure out 
how it can be set up. Undoubtedly anonymity is one of the main sources of deviant behaviour \citep{Knight:1999}.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.30\textwidth]{chap9/figs/eyes.pdf}
\includegraphics[width=.30\textwidth]{chap9/figs/identified-colors.pdf}}
\caption{\label{fig:eyes}Example interaction of the ``Look into the Box" art piece. 
Left: Top and pixelated eye of a spectator. Right: the main colours that 
were extracted from this image. Agents played language games to describe these eye colours.}
\end{figure}

### 2015-04-27 extra/papers-language-steels.tex
fixed remaining typos

9. \citet{Steels:97c} The Origins of Syntax in Visually Grounded Robotic Agents. In Pollack, M. (ed.) Proceedings of the 15th International Joint Conference on Artificial Intelligence, Morgan Kauffman Publishers, San Francisco, CA. pp. 1632-1641.

@@

39. \citet{Steels:01d} Linguistique evolutionaire et Vie Artificielle. Vie Artificielle, 1(1). Hermes Pub. Paris.

@@

55. \citet{Neubauer:03g}  Compositionality in Horizontal Transmission.  Poster European Cognitive Science Conference, 
University of Osnabruck.

>>>>

9. \citet{Steels:97c} The Origins of Syntax in Visually Grounded Robotic Agents. In Pollack, M. (ed.) Proceedings of the 15th International Joint Conference on Artificial Intelligence, Morgan Kaufmann Publishers, San Francisco, CA. pp. 1632-1641.

@@

39. \citet{Steels:01d} Linguistique evolutionnaire et Vie Artificielle. Vie Artificielle, 1(1). Hermes Pub. Paris.

@@

55. \citet{Neubauer:03g}  Compositionality in Horizontal Transmission.  Poster European Cognitive Science Conference, 
University of Osnabrück.
